---
title: "Final Project"
Author: ""
date: '`r Sys.Date()`'
editor_options: 
  chunk_output_type: console
output:
  pdf_document: default
  html_document: default
---

1. Introduction
Briefly introduce the research question and its significance. and Problem statement
Explain the purpose of the analysis: to understand factors influencing student lifestyle and expenses in major Saudi cities.
Provide a high-level overview of the dataset and the variables you will be working with.

```{r}
j
```


2. Loading and Exploring Data
Load the survey data into R from the appropriate file format (CSV or RData).
Display the first few rows of the dataset to get an overview of the data structure.
Check for missing values and handle them appropriately.
Explore summary statistics of key variables (mean monthly expenses, demographic information, spending categories) to gain initial insights.
```{r}

```


3. Data Preprocessing
Clean and preprocess the data as necessary (handling missing values, transforming variables, etc.).
Create dummy variables for categorical predictors if needed.
Normalize or scale continuous variables if required for the chosen modeling techniques.
```{r}

```


4. Exploratory Data Analysis (EDA)
Conduct exploratory data analysis using visualizations (histograms, box plots, etc.) to understand the distribution of variables.
Explore correlations between predictor variables and the outcome variable (monthly expenses).
Generate insights into potential patterns and relationships within the data.
```{r}

```

5. Model Selection and Justification
Choose appropriate machine learning models for prediction (e.g., linear regression, random forest, etc.).
Justify your choice of models based on the nature of the data and the research question.
Split the dataset into training and testing sets for model validation.
```{r}

```


6. Model Training and Evaluation
Train the selected models using the training dataset.
Evaluate model performance using appropriate metrics (e.g., RMSE, R-squared) on the testing dataset.
Compare the performance of different models if multiple models are used.
```{r}

```


7. Interpretation of Results
Interpret the coefficients (if applicable) of the predictors in the selected model(s).
Discuss the significance of predictors and how they relate to students' spending patterns.
Use interpretable machine learning techniques (Partial Dependence Plots, Surrogate Models) to explain the impact of key variables.
```{r}

```


8. Discussion and Conclusion
Summarize the key findings from the analysis.
Discuss the implications of the results in the context of the research question.
Address limitations and potential sources of bias in the analysis.
Provide insights into how the findings can inform policies, support systems, and business strategies for students in urban Saudi environments.
```{r}

```

9. Future Work
Suggest potential areas for future research or analysis related to student spending patterns.
Discuss any additional data or variables that could enhance the analysis if available.
```{r}

```


10. References
Include references to relevant literature, datasets, and tools used in the analysis.
Remember to include well-commented R code throughout the document to explain each step of the analysis clearly. This structure will help you organize your R Markdown file systematically and present your findings coherently. Good luck with your analysis!
```{r}

```