---
title: "Predicting and Analyzing College Student Lifestyle and Spending Patterns in Major Cities of Saudi Arabia"
output: pdf_document
date: "2023-11-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

o What is the aim of the project?
– Summarize the problem
– State your goals
o What do you do in this report?
– offer a roadmap of the project


# Problem Statement and Background
o Give a clear and complete statement of the problem and/or aim of your
analysis.
o Include a brief summary of any related work that has tried to tackle a project
similar to yours (i.e. a light literature review)


# Data
o What is the unit of observation?
o What is the outcome variable?
– How is it measured?
– Where does it come from?
– Please describe how the outcome variable is distributed using a table
and/or graph.
o What are your predictor variables?
– How are they measured?
– Where do they come from?
– Please describe how the predictor variable are distributed using a table
and/or graph.
o Outline any potential issues with the data:
– Missingness
– Lack of variation and/or availability
– Any potential sources of bias
o How do you overcome/mitigate these issues in your analysis?

# Methodology


# Analysis
o Describe the methods/tools you explored in your project.
o Outline in detail our entire analysis.
– Justify the tools/methods that you used.
– Assume the reader is smart but doesn’t know R/Machine Learning well. That
is, be crystal clear about what you’re doing and why.
o Note that this section should walk us through what you’re doing and how you’re
planning on doing it. There should be no results presented in this section

# Results
o Give a detailed summary of your results. Present your results clearly and
concisely.
o Please use visualizations and tables whenever possible.
o Be sure to:
– Discuss the performance of your predictive model.
– Use interpretable machine learning to talk about which variables were
important in the prediction task and how they relate to the outcome (i.e.
PDP/ICE/Surrogate Models)

# Discussion
o What conclusions should we pull from your analysis?
o What are the limitations (i.e. what can’t we say given your findings)?
o How would you expand the analysis if given more time?
o Speak on the “success” of your project (as defined in your proposal).
– Did you achieve what you set out to do? If not why?

# References 
# Appendix 





```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
